The regularity of ODEs and thimble integrals \\ with respect to Borel summation

### 2024-03-04 exp-resurgence.tex
Argue that special ODE solutions have asymptotic expansions

% regular singular volterra operator
\newcommand{\volterra}{\mathcal{V}}
\newcommand{\hardpart}{\mathcal{V}_0}
\newcommand{\softpart}{\mathcal{V}_\star}
%%\newcommand{\kerwhole}{k}
\newcommand{\hardker}{k_0}
%%\newcommand{\softker}{k_\star}
\newcommand{\solwhole}{f}
\newcommand{\solproto}{f_0}
\newcommand{\solptb}{f_\star}
\newcommand{\fmlsolwhole}{\series{f}}
\newcommand{\fmlsolproto}{\series{f}_0}
\newcommand{\fmlsolptb}{\series{f}_\star}
%%\newcommand{\roots}{\mathfrak{B}}

@@

%We can use the Laplace transform to convert $\mathcal{P}$ into a family of integral operators $\hat{\mathcal{P}}_\upsilon$ parameterized by the point $\zeta = \upsilon$ on the position domain.
\color{Maroon}
\begin{theorem}\label{thm:borel_reg_ODE}
Choose a root $-\alpha$ of $P$. Choose an open sector $\Omega_\alpha$ which has an opening angle of $\pi$ or less, has $\zeta = \alpha$ at its tip, and doesn't touch any other root of $P(-\zeta)$. Equation $\mathcal{P}\Psi=0$ has a unique solution $\Psi_\alpha$ in the affine subspace
\[ z^{-\tau_\alpha} + \dualsingexpalg{-\tau_\alpha-1}(\Omega_\alpha) \]
of the space $\dualsingexpalg{-\tau_\alpha}(\Omega_\alpha)$ from Section~\ref{sec:reg-decay}. That solution is Borel regular.
\end{theorem}
\textcolor{violet}{If you prefer to work in the analytic world, you can use this result to prove the existence and Borel summability of special formal transmonomial solutions.}
\begin{corollary}\label{cor:borel_reg_ODE}
For each root $-\alpha$ of $P$, equation $\mathcal{P}\Psi=0$ has a formal solution in the transmonomial space $e^{-\alpha z} z^{-\tau_\alpha}\,\C \llbracket z^{-1} \rrbracket$, which is unique up to scaling. That solution is Borel summable, and its Borel sum is a constant multiple of $\Psi_\alpha$.
\end{corollary}
\textcolor{violet}{On the other hand, if you prefer to work in the formal world, you can use Theorem~\ref{thm:borel_reg_ODE} to show that the Poincar\'{e} solutions are Borel summable. See Section~\ref{sec:new-summability-proof}.}
\textcolor{violet}{[In another section, do the ``mirror image'' argument. Use the Poincar\'{e} algorithm to construct a formal transmonomial solution in the classical way. Use the Ramis Index Theorem to show that it's Gevrey-$1$, and then then use our existence and uniqueness result for regular singular Volterra equations to show that it's not only Gevrey-$1$, but in fact Borel summable.]}
%%The formal solution whose existence is shown in Corollary~\ref{cor:borel_reg_ODE} is the Poincar\'{e} solution.
%%\begin{corollary}
%%\textcolor{orange}{[The Poincar\'{e} solution is Borel summable.]}
%%\end{corollary}
\color{black}
\textcolor{Violet}{[This is sort of a mirror image of the Ramis Index Theorem?]}
%\begin{theorem}[see Theorem \ref{thm2-dim}]\label{thm2}
%The formal frame of solutions $\tilde{\Psi}_j$ is Borel summable \textcolor{orange}{[summability is the strongest part of the result]} and its Borel sum is proportional to $\Psi_j$. In fact, $\Psi_j$ is a Borel regular solution of \eqref{eqn:standard ODE}. \textcolor{orange}{[Alternate statement: look in Sauzin or Delabaere to get uniqueness of formal solution.]} \textcolor{orange}{[Split into theorem, which shows existence of Borel regular frame, and corollary, which shows that this implies a certain frame of formal solutions is Borel-summable and sums to actual solutions.]}
%\end{theorem}
%
\textcolor{red}{[Change or actually delete]}We summarize the main steps of the proof of Theorem~\ref{thm:borel_reg_ODE} and Corollary~\ref{cor:borel_reg_ODE} (the rigorous proofs are given in Section~\ref{borel_reg-ODE}). On the right-hand side of the diagram, we have the formal solution $\series{\Psi }_\alpha\in e^{-\alpha z}z^{-\tau_\alpha}\C\llbracket z^{-1}\rrbracket$ and its Borel transform $\series{\psi}_\alpha$. %Then, the theory of $1$-summability allows us to draw the ``square'' with corners $\hat{\psi}_j$ and its Laplace transform $\hat{\Psi}_j$.

@@

@@

  %arc (160:112:\spill) -- (112:2) .. controls (112:1.7) and (103:1.7) .. (103:2) -- (103:\spill)
  %arc (103:50:\spill) -- (50:2.6) .. controls (50:2.2) and (43:2.2) .. (43:2.6) -- (43:\spill)
  %arc (43:-52:\spill) -- (-52:2.3) .. controls (-52:2) and (-60:2) .. (-60:2.3) -- (-60:\spill)
  %arc (-60:-130:\spill) -- (0, 0);
  %%arc (46:-51:\spill) -- (-51:2.5) .. controls (-51:6.4) and (-55.8:6.4) .. (-55.8:-1) -- (-55.8:\spill);
 %arc (-59:-57:\spill)--(0, 0);

@@

%%\node[anchor=south east, outer sep=1mm] at (46.5:2.46) {$\zeta=\beta$};

@@

  %\draw[dashed,violet!50] \p -- (0,0);

>>>>

% regular singular volterra operator
\newcommand{\volterra}{\mathcal{V}}
\newcommand{\hardpart}{\mathcal{V}_0}
\newcommand{\softpart}{\mathcal{V}_\star}
%%\newcommand{\kerwhole}{k}
\newcommand{\hardker}{k_0}
%%\newcommand{\softker}{k_\star}
\newcommand{\solwhole}{f}
\newcommand{\solproto}{f_0}
\newcommand{\solptb}{f_\star}
\newcommand{\fmlsolwhole}{\series{f}}
\newcommand{\fmlsolproto}{\series{f}_0}
\newcommand{\fmlsolptb}{\series{f}_\star}
%%\newcommand{\roots}{\mathfrak{B}}
\DeclareMathOperator{\hol}{hol}

@@

%We can use the Laplace transform to convert $\mathcal{P}$ into a family of integral operators $\hat{\mathcal{P}}_\upsilon$ parameterized by the point $\zeta = \upsilon$ on the position domain.
\color{BlueViolet}
\begin{theorem}\label{thm:exist_uniq_ODE}
Choose a root $-\alpha$ of $P$. Choose an open sector $\Omega_\alpha$ which has an opening angle of $\pi$ or less, has $\zeta = \alpha$ at its tip, and doesn't touch any other root of $P(-\zeta)$. The equation $\mathcal{P}\Psi = 0$ has a unique solution $\Psi_\alpha$ in the affine subspace
\begin{theorem}\label{thm:has_ae_ODE}
The solution $\Psi_\alpha$ from Theorem~\ref{thm:exist_uniq_ODE} has an asymptotic expansion in the transmonomial space $e^{-\alpha z} z^{-\tau_\alpha}\,\C \llbracket z^{-1} \rrbracket$. In fact, up to scaling, it's asymptotic to the Poincar\'{e} solution found in that space \textcolor{orange}{[not defined yet?]}.
Combining our uniqueness and regularity results for $\psi_\alpha$ with uniqueness and regularity results for the Poincar\'{e} solutions, we can show that $\Psi_\alpha$ is Borel regular.
\begin{corollary}
The solution $\Psi_\alpha$ from Theorem~\ref{thm:exist_uniq_ODE} is Borel regular.
\end{corollary}
\begin{theorem}\label{thm:borel_reg_ODE}
Choose a root $-\alpha$ of $P$. Choose an open sector $\Omega_\alpha$ which has an opening angle of $\pi$ or less, has $\zeta = \alpha$ at its tip, and doesn't touch any other root of $P(-\zeta)$. Equation $\mathcal{P}\Psi=0$ has a unique solution $\Psi_\alpha$ in the affine subspace
of the space $\dualsingexpalg{-\tau_\alpha}(\Omega_\alpha)$ from Section~\ref{sec:reg-decay}. That solution is Borel regular.
\textcolor{violet}{If you prefer to work in the analytic world, you can use this result to prove the existence and Borel summability of special formal transmonomial solutions.}
\begin{corollary}\label{cor:borel_reg_ODE}
For each root $-\alpha$ of $P$, equation $\mathcal{P}\Psi=0$ has a formal solution in the transmonomial space $e^{-\alpha z} z^{-\tau_\alpha}\,\C \llbracket z^{-1} \rrbracket$, which is unique up to scaling. That solution is Borel summable, and its Borel sum is a constant multiple of $\Psi_\alpha$.
\end{corollary}
\textcolor{violet}{On the other hand, if you prefer to work in the formal world, you can use Theorem~\ref{thm:borel_reg_ODE} to show that the Poincar\'{e} solutions are Borel summable. See Section~\ref{sec:new-summability-proof}.}
\textcolor{violet}{[In another section, do the ``mirror image'' argument. Use the Poincar\'{e} algorithm to construct a formal transmonomial solution in the classical way. Use the Ramis Index Theorem to show that it's Gevrey-$1$, and then then use our existence and uniqueness result for regular singular Volterra equations to show that it's not only Gevrey-$1$, but in fact Borel summable.]}
%%The formal solution whose existence is shown in Corollary~\ref{cor:borel_reg_ODE} is the Poincar\'{e} solution.
%%\begin{corollary}
%%\textcolor{orange}{[The Poincar\'{e} solution is Borel summable.]}
%%\end{corollary}
\textcolor{Violet}{[This is sort of a mirror image of the Ramis Index Theorem?]}
%\begin{theorem}[see Theorem \ref{thm2-dim}]\label{thm2}
%The formal frame of solutions $\tilde{\Psi}_j$ is Borel summable \textcolor{orange}{[summability is the strongest part of the result]} and its Borel sum is proportional to $\Psi_j$. In fact, $\Psi_j$ is a Borel regular solution of \eqref{eqn:standard ODE}. \textcolor{orange}{[Alternate statement: look in Sauzin or Delabaere to get uniqueness of formal solution.]} \textcolor{orange}{[Split into theorem, which shows existence of Borel regular frame, and corollary, which shows that this implies a certain frame of formal solutions is Borel-summable and sums to actual solutions.]}
\textcolor{red}{[Change or actually delete]}We summarize the main steps of the proof of Theorem~\ref{thm:borel_reg_ODE} and Corollary~\ref{cor:borel_reg_ODE} (the rigorous proofs are given in Section~\ref{borel_reg-ODE}). On the right-hand side of the diagram, we have the formal solution $\series{\Psi }_\alpha\in e^{-\alpha z}z^{-\tau_\alpha}\C\llbracket z^{-1}\rrbracket$ and its Borel transform $\series{\psi}_\alpha$. %Then, the theory of $1$-summability allows us to draw the ``square'' with corners $\hat{\psi}_j$ and its Laplace transform $\hat{\Psi}_j$.

@@

\color{BlueViolet}
\begin{theorem}[restatement of \ref{thm:exist_uniq_ODE}]\label{thm:exist_uniq_ODE_proof}
Choose a root $-\alpha$ of $P$. Choose an open sector $\Omega_\alpha$ which has an opening angle of $\pi$ or less, has $\zeta = \alpha$ at its tip, and doesn't touch any other root of $P(-\zeta)$. The equation $\mathcal{P}\Psi = 0$ has a unique solution $\Psi_\alpha$ in the affine subspace
After rescaling by a constant, Theorem~4 of \cite{reg-sing-volterra} tells us that the equation $\hat{\mathcal{P}}_\alpha \psi_\alpha = 0$ has a unique solution $\psi_\alpha$ in the affine subspace
\textcolor{orange}{[\ldots]} Our arguments in \cite{reg-sing-volterra} took place on $\C$, but they work work just as well on a covering of $\C$ which is branched at the roots of $P(-\zeta)$. Let's use the covering $\pi \maps \widetilde{B}_{P(-\zeta)} \to \C$ built by puncturing $\C$ at the roots of $P(-\zeta)$, taking the universal cover, and then gluing the roots back in as infinite-order branch points.
\begin{lemma}\label{lem:shifted_holo_closed}
Choose a root $-\alpha$ of $P$. Consider a simply connected open set $\Upsilon \subset \widetilde{B}_{P(-\zeta)}$ which touches but doesn't contain $\zeta = \alpha$, is star-shaped around $\zeta = \alpha$, and doesn't touch any other root of $P(-\zeta)$. Suppose that $\pi \maps \Upsilon \to \C$ overlaps itself at the edges: it's two-to-one over a simply connected open set $\Upsilon_\Delta \subset \C$ which is star-shaped around $\zeta = \alpha$, and it's one-to-one everywhere else.
    (190:1.5) arc (190:41:1.5) -- (41:\spill) arc (41:-48:\spill) -- (-48:1.3) arc (-48:-170:1.3) -- cycle;
  \fill[violet!35] (0, 0) -- (-170:1.3) arc (-170:-210:1.3) -- cycle;
  \fill[violet] circle (0.7mm) node[anchor=west, outer sep=1mm] {$\zeta = \alpha$};
  \node[violet] at (-190:0.85) {$\Upsilon_\Delta$};
For any real numbers $\sigma$, $\Lambda$ and any $\delta \in [0, 1)$, the intersection of $\zeta^{\sigma + \delta}\,\mathcal{O}(\pi \Upsilon)$ with $\singexp{\sigma}{\Lambda}(\Upsilon)$ is a closed subspace of the latter.
Each point $b \in \Upsilon_\Delta$ has two preimages $b_+, b_- \in \Upsilon$, where $b_+$ is $2\pi$ counterclockwise of $b_-$ around $\zeta = \alpha$. The operator $\hol \maps \singexp{\sigma}{\Lambda}(\Upsilon) \to \singexp{\sigma}{\Lambda}(\Upsilon_\Delta)$ defined by
\[ \hol \varphi \big|_b = \varphi(b_+) - \varphi(b_-) \]
is bounded. The intersection of $\zeta^{\sigma + \delta}\,\mathcal{O}(\Upsilon_\Delta)$ with $\singexp{\sigma}{\Lambda}(\Upsilon)$ is the kernel of the bounded operator $\hol - e^{2\pi i \delta}$, so it's closed.
\begin{theorem}[restatement of \ref{thm:has_ae_ODE}]
The solution $\Psi_\alpha$ from Theorem~\ref{thm:exist_uniq_ODE_proof} has an asymptotic expansion in the transmonomial space $e^{-\alpha z} z^{-\tau_\alpha}\,\C \llbracket z^{-1} \rrbracket$. In fact, up to scaling, it's asymptotic to the Poincar\'{e} solution found in that space \textcolor{orange}{[not defined yet?]}.
Take an open sector $\Omega_\alpha$ of the kind used in Theorem~\ref{thm:exist_uniq_ODE_proof}, and expand it to an open set $\Upsilon \subset \widetilde{B}_{P(-\zeta)}$ of the kind used in Lemma~\ref{lem:shifted_holo_closed}. By Theorem~4 of \cite{reg-sing-volterra}, the equation $\hat{\mathcal{P}}_\alpha \psi_\alpha = 0$ has a unique solution $\psi_\alpha$ in the affine subspace
\[ \frac{\zeta^{\tau_\alpha-1}}{\Gamma(\tau_\alpha)} + \singexpalg{\tau_\alpha}(\Upsilon) \]
of the space $\singexpalg{\tau_\alpha-1}(\Upsilon)$. Restricting this solution to $\Omega_\alpha$ gives the analogous solution from the proof of Theorem~\ref{thm:exist_uniq_ODE_proof}, as we can see from the uniqueness part of Theorem~4. To prove that $\Psi_\alpha$ has a the desired asymptotic expansion, we'll show that $\psi_\alpha$ lies in $\zeta^{\tau_\alpha - 1}\,\mathcal{O}(\pi \Upsilon)$.

%%The key step of our proof will be to show that $\psi_\alpha$ lies in

%%The uniqueness part of the Theorem~4 ensures that restricting $\psi_\alpha$ to $\Omega_\alpha$ gives the analogous solution used in the proof of Theorem~\ref{thm:exist_uniq_ODE_proof}.

%%%In the proof of Theorem~\ref{thm:exist_uniq_ODE_proof}, we used Theorem~4 of \cite{reg-sing-volterra} to show that the equation $\mathcal

%%%$\psi_\alpha$ in the affine subspace
%%\[ \frac{\zeta^{\tau_\alpha-1}}{\Gamma(\tau_\alpha)} + \singexpalg{\tau_\alpha}(\Omega_\alpha) \]
%%of the space $\singexpalg{\tau_\alpha-1}(\Omega_\alpha)$

For convenience, let $p = P(-\zeta)$ and $q = Q(-\zeta)$. In \cite{reg-sing-volterra}, to prove Theorem~4, we rewrite the equation $\hat{\mathcal{P}}_\alpha \solwhole = 0$ as a regular singular Volterra equation $\solwhole = \volterra^\alpha \solwhole$ that satisfies the conditions of Theorem~3. The operator $\volterra^\alpha$ is the sum of the ``prototypical'' part
%%\[ \hardpart^\alpha = -\tfrac{1}{P(-\zeta)} \circ \partial^{-1}_{\zeta, \alpha} \circ Q(-\zeta) \]
\[ \hardpart^\alpha = -\tfrac{1}{p} \circ \partial^{-1}_{\zeta, \alpha} \circ q \]
and the perturbation
%%\[ \softpart^\alpha = \tfrac{1}{P(-\zeta)} \circ \partial^{-2}_{\zeta, \alpha} \circ R(\partial^{-1}_{\zeta, \alpha}) \]
\[ \softpart^\alpha = \tfrac{1}{p} \circ \partial^{-2}_{\zeta, \alpha} \circ R\big(\partial^{-1}_{\zeta, \alpha}\big). \]
\color{SlateBlue}
\begin{itemize}
\item Prove Theorem~3 using Picard iteration, using the {\em prototype solution}
\[ \solproto(a) = \frac{1}{p(a)} \left(\frac{\zeta(a)}{\zeta(b)}\right)^\tau \exp\left[-\int_b^a \left( \frac{q}{p} + \frac{\tau}{\zeta} \right)\;d\zeta\right], \]
given by an arbitary point $b \in \Omega_\alpha$ as our initial guess.
\item This expression for the prototype solution, taken from \textcolor{orange}{\tt sec:construction} Section~3.2.1 of \cite{reg-sing-volterra}, makes it clear that the prototype solution belongs to $\zeta^{\tau_\alpha - 1}\,\mathcal{O}(\pi \Upsilon)$.
\item The operator $\volterra^\alpha$ preserves the intersection of $\zeta^{\sigma + \delta}\,\mathcal{O}(\pi \Upsilon)$ with $\singexp{\sigma}{\Lambda}(\Upsilon)$. Hence, the Picard iteration sequence must lie in this subspace.
\item We know from Lemma~\ref{lem:shifted_holo_closed} that the intersection of $\zeta^{\tau_\alpha - 1}\,\mathcal{O}(\pi \Upsilon)$ with $\singexp{\sigma}{\Lambda}(\Upsilon)$ is a closed subspace of the latter. Hence, the solution $\psi_\alpha$ that Picard iteration converges to must also lie within this subspace.
\item Use Laplace transforms of truncated power series for $\psi_\alpha$ to show that $\Psi_\alpha$ has an asymptotic series?
\end{itemize}
\color{BlueViolet}

%%Let's run through the proof of Theorem~3, specialized to the case we consider in Theorem~4. First, picking an arbitrary point $b \in \Omega_\alpha$, we show that

%%Picking an arbitrary point $b \in \Omega_\alpha$, we observe that the {\em prototype solution}
%%\[ \solproto(a) = \frac{1}{p(a)} \left(\frac{\zeta(a)}{\zeta(b)}\right)^\tau \exp\left[-\int_b^a \left( \frac{q}{p} + \frac{\tau}{\zeta} \right)\;d\zeta\right] \]

%% \left(\frac{\zeta(a)}{\zeta(b)}\right)^\tau \exp\left[-\int_b^a \left( \frac{q}{p} + \frac{\tau}{\zeta} \right)\;d\zeta\right]

%%Theorem~4 of \cite{reg-sing-volterra} tells us that the equation $\hat{\mathcal{P}}_\alpha \psi_\alpha = 0$ has a unique solution $\psi_\alpha$ in the subspace

%%The restriction map $\singexp{\sigma}{\Lambda}(\Upsilon) \hookrightarrow \singexp{\sigma}{\Lambda}(\Omega_\alpha)$ \textcolor{orange}{\ldots}

@@

@@

@@
